### 2024-04-27 article.tex
MFCC explanation added

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Graphics/naturalist_horizontal.png}
			\caption{Image file count for each animal}
			\label{fig:enter-label}
		\end{figure}

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Graphics/xenocanto.png}
			\caption{Image file count for each bird}
			\label{fig:enter-label}
		\end{figure}

	}
\section{Proposed Methodlogy}
\subsection{Mel-frequency cepstrum Coefficients based on DCT-II for Audio Feature Extraction
}
Mel-frequency cepstrum (MFC) refers to the short-term power spectrum of a sound
based on cosine transform. Mel-frequency cepstrum Coefficients(MFCCs) are coefficients that collectively make up an MFC. Mel-frequency cepstrum Coefficients transform is done on the audio to extract a small set of features that describes our audio.
MFCC helps us to capture the essential spectral characteristics of speech signals in a
compact form, making them suitable for pattern recognition tasks.
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
def mfcc_transform(y: np.ndarray,
                   sr: int,
                   mfcc_params: Dict
                   ) -> torch.Tensor:
    melspec = librosa.feature.melspectrogram(
        y=y,
        sr=sr,
        n_fft=mfcc_params['n_fft'],
        win_length=mfcc_params['win_length'],
        hop_length=mfcc_params['hop_length'],
        n_mels=mfcc_params['n_mels'],
        htk=mfcc_params['htk'],
        norm=mfcc_params['norm_melspec'],
    )
    mfcc_librosa = librosa.feature.mfcc(
        S=librosa.core.spectrum.power_to_db(melspec),
        n_mfcc=mfcc_params['n_mfcc'],
        dct_type=mfcc_params['dct_type'],
        norm=mfcc_params['norm_mfcc']
    )

    return torch.from_numpy(mfcc_librosa)
\end{lstlisting}

\subsection{Expansion and Interpolation of Audio Data}
After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
\begin{itemize}
	\item The input tensor y is replicated along one dimension to create three copies, preparing it for subsequent operations.
	\item The expanded tensor is resized to a fixed size of (224, 224) using bilinear interpolation, ensuring consistent dimensions for neural network input.
	\item Any extra dimensions introduced during resizing are removed, returning the tensor to its original shape.
	\item The processed tensor is returned for further use, completing the preprocessing steps commonly applied to image data for deep learning tasks.
\end{itemize}
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
def expand_tensor_for_efficient_net(y) -> torch.Tensor:
    input_tensor_expanded = y.expand(3, -1, -1)
    resized_tensor = F.interpolate(input_tensor_expanded.unsqueeze(
        0), size=(224, 224), mode='bilinear', align_corners=False)

    resized_tensor = resized_tensor.squeeze()
    return resized_tensor
\end{lstlisting}

\subsection{Transformation as required by EfficientNetB0 Input Data}
For best results and compatibility, we transformed our data as per the transformation required for EfficientNetB0 architecture.
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
ImageClassification(
    crop_size=[224]
    resize_size=[256]
    mean=[0.485, 0.456, 0.406]
    std=[0.229, 0.224, 0.225]
    interpolation=InterpolationMode.BICUBIC
)
\end{lstlisting}

@@

\subsection{EfficientNetB0}
EfficientNet is a convolutional neural network architecture and scaling method that uni-
formly scales all dimensions of depth/width/resolution using a compound coefficient \cite{r7}.
Unlike conventional practice that arbitrarily scales these factors, the EfficientNet scaling method uniformly scales network width, depth, and resolution with a set of fixed scaling
coefficients.
\par
\subsection{ Convolutional Neural Network}
Convolutional Neural Networks are specialized kind of neural networks for processing
data that has a grid-like topology \cite{r8}. Most machine learning libraries implement cross-correlation rather than conventionally known convolution operations in the realm of Digital
Signal Processing . In conventional multi-layer perceptrons, large number of parameters are required to learn features where as CNN uses filters to stride through whole data using a lesser number of parameters aka parameter sharing. CNNs are often applied
in large number of computer vision tasks across the industry.

\section{Result and Analysis}

\begin{center}
	\captionof{table}{Overview of Results of Audio and Video Model}
	\begin{tabular}{ |c|c|c| }
		\hline
		Model  & Accuracy & F1 score \\
		\hline
		Vision & 86.1     & 86.12    \\
		\hline
		Audio  & 90.39    & 88.12    \\

>>>>

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Graphics/naturalist_horizontal.png}
			\caption{Image file count for each animal}
			\label{fig: image_file_count}
		\end{figure}

@@

		\begin{figure}
			\centering
			\includegraphics[scale=0.5]{Graphics/xenocanto.png}
			\caption{Audio file count for each bird}
			\label{fig: audio_file_count}
		\end{figure}

	}
\section{Proposed Methodlogy}
\subsection{Extraction of Frames and Audio from Video}
The user can upload video files of different durations. To create proper input for our audio and vision model, first the audio is extracted from the video which is then broken down into 5 sec chunks. Moreover, the image frames are extracted from the video at the rate of 1fps. Both audio and video are seperately sent to the audio and vision model respectively which then returns the output which is displayed to the user.
The following steps are involved in temporal segmentation of audio after extraction:
\begin{itemize}
	\item {The duration of the audio is calculated.}
	      \item{The number of 5 second audio segments that can be created is calculated. }
	\item { The audio is then split into 5 second segments. If any of the segments is less than 5 seconds then they are discarded. }
\end{itemize}
The following steps are involved in the splitting of video into image frames:
\begin{itemize}
	\item {The duration of video is calculated.}
	      \item{At the rate of 1fps, image frames are extracted from the video using ffmpeg.}
\end{itemize}
\subsection{Mel-frequency cepstrum Coefficients based on DCT-II for Audio Feature Extraction
}
Mel-frequency cepstral coefficients (MFCCs) are a set of features which represents the characteristics of sound. These coefficients are generated in two steps. First, a mel-scaled spectogram is created from the audio which are then used to generate MFCCs.
Mel Spectogram is a graphic representation of a sound wave. It visualises frequency over time. The process of generating a Mel Spectogram can be explained in the following steps:
\begin{itemize}
	\item {The audio signal is broken down into short frames.}
	\item {The time signal is converted into frequency domain using a STFT.}
	      \item{The frequencies are then converted into Mel scale.}
\end{itemize}
The mel scale is a perceptual scale of pitches judged by listeners to be equal in distance from one another. The following formulae is used to convert frequencies to mel scale.
\[
	Mel(f)=1125ln(1+ \frac{f}{100})
\]
The values computed above are highly correlated, which is not ideal for machine learning tasks. To solve this, Discrete Cosine Transform-II (DCT-II) is applied to decorrelate the coefficients. The then generated coefficients are also normalized orthogonally. The choice of DCT II is primarily because of its ability to compactly represent signal energy. The number of MFCCs to return is chosen to be 256 because it was easier to later transform to appropriate size for EfficientNetB0.
\subsection{Expansion and Interpolation of Audio Data}
After MFCC transformation, the tensors are transformed again to proper dimension as required by EfficientNetB0 architecture. Following steps are performed:
\begin{itemize}
	\item The input tensor y is replicated along one dimension to create three copies, preparing it for subsequent operations.
	\item The expanded tensor is resized to a fixed size of (224, 224) using bilinear interpolation, ensuring consistent dimensions for neural network input.
	\item Any extra dimensions introduced during resizing are removed, returning the tensor to its original shape.
	\item The processed tensor is returned for further use, completing the preprocessing steps commonly applied to image data for deep learning tasks.
\end{itemize}
% \begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
% def expand_tensor_for_efficient_net(y) -> torch.Tensor:
%     input_tensor_expanded = y.expand(3, -1, -1)
%     resized_tensor = F.interpolate(input_tensor_expanded.unsqueeze(
%         0), size=(224, 224), mode='bilinear', align_corners=False)

%     resized_tensor = resized_tensor.squeeze()
%     return resized_tensor
% \end{lstlisting}

\subsection{Transformation as required by EfficientNetB0 Input Data}
For best results and compatibility, we transformed our data as per the transformation required for EfficientNetB0 architecture.
\begin{lstlisting}[style=mystyle,caption={MFCC Transformation Function},captionpos=b]
ImageClassification(
    crop_size=[224]
    resize_size=[256]
    mean=[0.485, 0.456, 0.406]
    std=[0.229, 0.224, 0.225]
    interpolation=InterpolationMode.BICUBIC
)
\end{lstlisting}
\subsection{ Convolutional Neural Network}
Convolutional Neural Networks are specialized kind of neural networks for processing
data that has a grid-like topology \cite{r8}. Most machine learning libraries implement cross-correlation rather than conventionally known convolution operations in the realm of Digital
Signal Processing . In conventional multi-layer perceptrons, large number of parameters are required to learn features where as CNN uses filters to stride through whole data using a lesser number of parameters aka parameter sharing. CNNs are often applied
in large number of computer vision tasks across the industry.

@@

\subsection{EfficientNetB0}
EfficientNet is a convolutional neural network architecture and scaling method that uni-
formly scales all dimensions of depth/width/resolution using a compound coefficient \cite{r7}.
Unlike conventional practice that arbitrarily scales these factors, the EfficientNet scaling method uniformly scales network width, depth, and resolution with a set of fixed scaling
coefficients.
\par
\section{Result and Analysis}

\begin{center}
	\captionof{table}{Overview of Results of Audio and Video Model}
	\begin{tabular}{ |c|c|c|c| }
		\hline
		Model  & Train Accuracy & Test Accuracy & F1 score \\
		\hline
		Vision & 96.00          & 86.10         & 86.12    \\
		\hline
		Audio  & 98.36          & 90.39         & 88.12    \\
